Efficient pedigree recording for fast population genetics simulation

### 2017-12-27 forwards_paper.tex
add missing words

However, it is thought that most genetic variation is selectively neutral (or nearly so).
By definition, the alleles carried by individuals in a population at neutral sites
do not affect the population process.
For this reason, if one records the entire genealogical history of a population over the course of a simulation,
simply laying down neutral mutations on top of that history afterwards
is equivalent to having generated them during the simulation.
Precisely, we would need to know the genealogical tree relating all sampled individuals
at each position along the genome.
In this paper, we describe how algorithmic tools and data structures developed for the
coalescent simulator \msprime{}
can be used to efficiently record, and later process, this history.
In so doing we record the \emph{population pedigree} --
the complete history of parent-offspring relationships of an entire population
going back to a remote time including the genetic outcomes of each ancestral
meiosis.
This embellished pedigree contains all the information necessary
to construct the genealogical tree that relates each individual to each other
at each position on the genome, i.e., the \emph{tree sequence}~\citep{kelleher2016efficient}.
Combined with ancestral genotypes and the origins of new mutations,
it also completely specifies the genomic sequence of any individual in the population at any time.
This is much more than we need to know, however,
so we discard all information irrelevant to the genetic history
of the extant population. We show that we need only $O(N \log N)$
space to store the resulting tree sequence for a single crossover per
generation Wright-Fisher model.

>>>>

However, it is thought that most genetic variation is selectively neutral (or nearly so).
By definition, the neutral alleles carried by individuals in a population
do not affect the population process.
For this reason, if one records the entire genealogical history of a population over the course of a simulation,
simply laying down neutral mutations on top of that history afterwards
is equivalent to having generated them during the simulation.
Precisely, we would need to know the genealogical tree relating all sampled individuals
at each position along the genome.
In this paper, we describe how algorithmic tools and data structures developed for the
coalescent simulator \msprime{}
can be used to efficiently record, and later process, this history.
In so doing we record the \emph{population pedigree} --
the complete history of parent-offspring relationships of an entire population
going back to a remote time including the genetic outcomes of each ancestral
meiosis.
This embellished pedigree contains all the information necessary
to construct the genealogical tree that relates each individual to each other
at each position on the genome, i.e., the \emph{tree sequence}~\citep{kelleher2016efficient}.
Combined with ancestral genotypes and the origins of new mutations,
it also completely specifies the genomic sequence of any individual in the population at any time.
This is much more than we need to know, however,
so we discard all information irrelevant to the genetic history
of the extant population. We show that we need only $O(N \log N)$
space to store the resulting tree sequence for a single crossover per
generation under the Wright-Fisher model.
